---
title: "data-wrangling"
author: "Maya Sollen-Norrlin"
date: "2022-08-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Loading libraries   
```{r}

library(tidyverse)
library(gridExtra)
library(ggpubr)
```


## Read in the data
```{r}

chem_data <- read.csv("chemistry-data-raw.csv", na.strings = c(""))

```

 

```{r}

str(chem_data)


```

```{r}
head(chem_data)

```
## Converting time variable into factor
```{r}

chem_data$Time..weeks. <- as.factor(chem_data$Time..weeks.)

```

```{r}

str(chem_data)

```
## Filter out irrelevant columns
```{r}

chem_data_filtered <- chem_data %>%
  select(Sample.ID, moisture...., pH, NO3..mg.kg.:NH4..mg.kg., K..mg.kg.:Time..weeks.)

str(chem_data_filtered)
```


```{r}
head(chem_data_filtered, 20)
```


```{r}

complete.cases(chem_data_filtered)

```

```{r}

tail(chem_data_filtered, 20)
```



```{r}


chem_data_filtered[!complete.cases(chem_data_filtered),]


```


Renaming columns so they're easier to write.
```{r}

str(chem_data_filtered)

```

```{r}
chem_data_filtered<- chem_data_filtered %>% 
  rename(
   "NH4 (mg/kg)" = NH4..mg.kg.,
   "K (mg/kg)" = K..mg.kg.,
   "NO2 (mg/kg)" = NO2..mg.kg.,
   "NO3 (mg/kg)" = NO3..mg.kg.,
   "Moisture (%)" = moisture....,
   "Weeks" = Time..weeks.
  )

head(chem_data_filtered)


```



## Replacing negative values in the moisture and nitrate columns with 0. 
```{r}

chem_data_filtered <- chem_data_filtered%>% 
  mutate(across(c(`Moisture (%)`, `NO3 (mg/kg)`:`NH4 (mg/kg)`), ~ ifelse(.x<0, 0, .x)))


tail(chem_data_filtered, 20)
```
## Save new data frame of tidy data for further analysis
```{r}

write_csv(chem_data_filtered, "chemistry-data-tidy.csv", col_names = TRUE, na = "")

```


Filtering out only the dry treatment data for calculating wet weights based on moisture
```{r}

dryData <- chem_data_filtered %>%
  filter(Treatment == "Dry")

```

Read in csv file with excact weights for each sample aliqout (these are dry weigths for the "Dry" category samples)
```{r}

weights <- read.csv("test-weights.csv")

```

Filter out only the weights for the Dry samples
```{r}

dryWeights <- weights %>%
  filter(Treatment == "Dry") %>%
  select(Sample.ID, X3g, X5g.K, X5g.NP, Treatment, Time..weeks. )

head(dryWeights)
```
Renaming columns
```{r}
dryWeights<- dryWeights %>% 
  rename(
   "3g" = X3g,
   "5g.K" = X5g.K,
   "5g.N" = X5g.NP,
   "Weeks" = Time..weeks.
  )

head(dryWeights)
```

```{r}
str(dryWeights)
```

```{r}

dryWeights$Weeks <- as.factor(dryWeights$Weeks)

str(dryWeights)
```
Combined the dry test weights with the rest of the data
```{r}

dryDataComb <- full_join(dryData, dryWeights, by = c("Sample.ID", "Treatment", "Weeks"))
str(dryDataComb)
```
Calculate the equivalent wet weight for each sample based on the lost moisture
```{r}

dryDataComb <- dplyr::mutate(dryDataComb, wetWeightK = ((`Moisture (%)`/100)*`5g.K`)+`5g.K`)

```

```{r}

dryDataComb <- dplyr::mutate(dryDataComb, wetWeightN = ((`Moisture (%)`/100)*`5g.N`)+`5g.N`)

```


Read in the raw nutrient results from the discrete analyser (in mg/l)
```{r}

rawNutrients <- read.csv("raw-nutrient-output.csv", na.strings = c(""))


str(rawNutrients)

```

```{r}

dryNutri <- rawNutrients %>%
  filter(Treatment == "Dry") 

str(dryNutri)
```

```{r}
dryNutri$Weeks <- as.factor(dryNutri$Weeks)

str(dryNutri)

```

```{r}
dryNutri<- dryNutri %>% 
 rename(
   "NH4 (mg/l)" = NH4..mg.l.,
   "K (mg/l)" = K..mg.l.,
   "NO2 (mg/l)" = NO2.as.N...mg.l.,
   "NO3 (mg/l)" = NO3.as.N..mg.l.,
  )

head(dryNutri)
```

```{r}

dryDataFinal <- full_join(dryDataComb, dryNutri, by = c("Sample.ID", "Treatment", "Weeks"))

dryDataFinal <- select(dryDataFinal, Sample.ID, Treatment, Weeks, wetWeightK:`K (mg/l)`)

dryDataFinal <- dryDataFinal %>%
                  filter(Weeks != "0")

str(dryDataFinal)
```


```{r}

dryDataFinal <- dplyr::mutate(dryDataFinal, soil.in.1L.N = wetWeightN*50)

dryDataFinal <- dplyr::mutate(dryDataFinal, soil.in.1L.K = wetWeightK*50)

dryDataFinal <- dplyr::mutate(dryDataFinal, dilutionFactorN = 1000/soil.in.1L.N)

dryDataFinal <- dplyr::mutate(dryDataFinal, dilutionFactorK = 1000/soil.in.1L.K)

dryDataFinal <- dplyr::mutate(dryDataFinal, `NO3 (mg/kg) dry` = `NO3 (mg/l)`*dilutionFactorN)

dryDataFinal <- dplyr::mutate(dryDataFinal, `NO2 (mg/kg) dry` = `NO2 (mg/l)`*dilutionFactorN)

dryDataFinal <- dplyr::mutate(dryDataFinal, `NH4 (mg/kg) dry` = `NH4 (mg/l)`*dilutionFactorN)

dryDataFinal <- dplyr::mutate(dryDataFinal, `K (mg/kg) dry` = `K (mg/l)`*dilutionFactorK)

str(dryDataFinal)
```

```{r}

dryDataFinal <- dryDataFinal %>% 
                  mutate_if(is.numeric, ~round(., 2))

```


```{r}

write_csv(dryDataFinal, "nutrients-dry-backCalculated.csv", col_names = TRUE, na = "")

```